---
title: "p8105_HW_2"
output: github_document
---

# Homework 2 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(knitr)
```

## Problem 1
### Creating Dataframe for Mr. Trash Wheel Data
```{r}
trash_wheel_df <- 
  read_excel("./data_hw_2/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(-c(x15, x16, x17)) %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls, 0) %>% 
      as.numeric(sports_balls)
  )
```

### Creating Dataframe for 2018 Precipitation
```{r}
precip_2018 <- 
  read_excel("./data_hw_2/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  drop_na(Total) %>% 
  mutate(
    year = 2018
  )
```

### Creating Dataframe for 2017 Precipitation
```{r}
precip_2017 <- 
  read_excel("./data_hw_2/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  drop_na(Total) %>% 
  mutate(
    year = 2017
  )
```

### Combining 2017 and 2018 Precipitation Datasets
```{r}
precip_17_18 <- 
  bind_rows(precip_2017, precip_2018) %>% 
  janitor::clean_names() %>% 
  mutate(
    month = month.name[month]
  )
```
Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017?

The Mr. Trash Wheel dataset is comprised of data that tracks the trash collected by Mr. Trash Wheel, with 'r ncol(trash_wheel_df)' variables and 'r nrow(trash_wheel_df)' observations. The month, day, and year is recorded along with how many dumpsters were filled. The volume and weight of the collected trash is accounted for. In addition, the type of trash collected is broken down into the following categories: plastic bottles, polystyrene, ciagarette butts, glass bottles, grocery bags, chip bags, and sports balls. There is also a "Homes Powered" variable, as the garbage collected is eventually incinerated to create electricity for homes. The median number of sports balls in a dumpster in 2017 is 'r '.

### Remember to fix median number of sports balls!

The Mr. Trash Wheel dataset is also comprised by sheets of annual precipitation data; in this case 2018 and 2017 precipitation data was looked at. For the 2018 precipitation data, there are 'r nrow(precip_2018)' observations for each month of the year and 'r ncol(precip_2018)' observations, Month, total, and year respectively. The totaly precipitation in 2018 was 'r sum(pull(precip_2018, Total))'. Similarly for the 2017 precipitation data, there are 'r nrow(precip_2017)' observations for each month of the year and 'r ncol(precip_2017)' observations, Month, total, and year respectively.


## Problem 2
### Tidying the dates 
```{r}
pols_month_df <- 
  read_csv("./data_hw_2/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, c("year", "month", "day")) %>%
  mutate(
    month = as.numeric(pull(., month)),
    month = month.name[month]
    )
```

### Create president variable
```{r}
pols_month_df <- 
  pivot_longer(
    pols_month_df,
    cols = c("prez_gop", "prez_dem"),
    names_to = "president",
    names_prefix = "prez_"
  ) %>% 
  mutate(value = na_if(value, 0)) %>% 
  drop_na() %>% 
  select(-value, -day)
```

### Cleaning snp.csv 
```{r}
snp_df <- 
  read_csv("./data_hw_2/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, c("month", "day", "year")) %>% 
  mutate(
    month = as.numeric(pull(., month)),
    month = month.name[month]
    ) %>% 
  select(year, month, close, -day) %>% 
  arrange(year, (match(month, month.name)))
```

### Cleaning Unemployment.csv
```{r}
unemploy_df <- read_csv("./data_hw_2/fivethirtyeight_datasets/unemployment.csv") %>% 
  rename(year = Year) %>% 
  mutate(
    year = as.character(year)
  )

unemploy_df <-
  pivot_longer(
    unemploy_df,
    cols = Jan:Dec,
    names_to = "month",
  ) %>% 
  mutate(
    month = match(month, month.abb),
    month = month.name[month],
  ) 
```

### Joining snp_df with pol_month_df and join_snp_pols_df with unemploy_df
```{r}
#Is this the way to do it?
join_snp_pols_df <- 
  left_join(pols_month_df, snp_df, by = c("year", "month"))

join_unemploy_df <- 
  left_join(join_snp_pols_df, unemploy_df, by = c("year", "month"))
```
Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

The 'r pols_month_df' refers to the dataset containing information regarding to the number of national politicians who are democratic or republican at a given time. There are 'r nrow(pols_month_df) observations and 'r ncol(pols_month_df) variables such as year, month, president (whether the president was a republican or democrat), sen_gop or sen_dem (how many senators were republican or democrat), rep_gop or rep_dem (how many house representations were republican or democrat), and gov_gop or gov_dem (how many governers were republican or democrat). This dataset's year variable ranges from 1947 to 2015. 

The 'r snp_df' refers to the dataset containing information related to the S&P market index, which is used as a representative measure of the stock market. There are 'r nrow(snp_df)' obserations and 'r ncol(snp_df)' variables such as year, month, and close (the closing values of the S&P stock index on the associated date). This dataset's year variable ranges from 1950 to 2015.

The 'r unemploy_df' refers to the dataset containing infomration related to the percentage of unemployment of the associated year. There are 'r nrow(unemploy_df)' observations and 'r ncol(unemploy_df)' variables such as year, month, and value (percentage of unemployment). This dataset's year variable ranges from 1948 to 2015. 

## Problem 3
### Cleaning the Popular_Baby_Names.csv 
```{r}
baby_names_df <- 
  read_csv("./data_hw_2/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct() %>% 
  arrange(year_of_birth) %>% 
  mutate_all(tolower) %>% 
  mutate(
    ethnicity = recode(ethnicity,
      "asian and paci" = "asian and pacific islander",
      "black non hisp" = "black non hispanic",
      "white non hisp" = "white non hispanic")
  )
```

### Creating a reader-friendly table showing the rank in popularity of the name "Olivia"
```{r results = 'asis'}
Oliva_tbl <-
  baby_names_df %>% 
  select(year_of_birth, ethnicity, childs_first_name, rank) %>% 
  filter(childs_first_name == "olivia") %>% 
  pivot_wider(
    names_from = year_of_birth, 
    values_from = rank,
  ) %>%
  kable()
```

### Table of the most popular male names over time
```{r}
male_names_tbl <-
  baby_names_df %>% 
  filter(gender == "male", rank == "1") %>% 
  select(year_of_birth, ethnicity, childs_first_name, rank) %>% 
  kable()
```

### Scatter plot showing number of children with a name against rank in popularity 
```{r}
white_hisp_male_df <- 
  baby_names_df %>% 
  filter(gender == "male", ethnicity == "white non hispanic", year_of_birth == "2016") %>% 
  mutate(
    rank = as.numeric(rank),
    count = as.numeric(count)
  ) 

ggplot(white_hisp_male_df, aes(x = rank, y = count)) +
  geom_point()
```

